High-Dimensional Regularized Discriminant Analysis

### 2013-05-12 reports/paper-hdrda.Rnw
Added supplementary material and keywords.

\begin{aug}
\author{\fnms{John A.} \snm{Ramey}\thanksref{m1}\ead[label=e1]{johnramey@gmail.com}}
\author{\fnms{Caleb K.} \snm{Stein}\thanksref{m2}\ead[label=e2]{Caleb\_Stein@baylor.edu}}
\and
\author{\fnms{Dean M.} \snm{Young}\thanksref{m2}\ead[label=e3]{Dean\_Young@baylor.edu}}

@@

\address{Vaccine and Infectious Disease Division\\
  Fred Hutchinson Cancer Research Center\\
  1100 Fairview Ave. N., M2-C200\\
  Seattle, Washington 98115, U.S.A.\\
  \printead{e1}}

@@

%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

%\begin{keyword}
%\kwd{}
%\kwd{}
%\end{keyword}

@@

% AOS,AOAS: If there are supplements please fill:
%\begin{supplement}[id=suppA]
%  \sname{Supplement A}
%  \stitle{Title}
%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
%  \sdatatype{.pdf}" 
%  \sdescription{Some text}
%\end{supplement}

@@

# Summary of the classification error rates for each data set and number of
# variables selected, d.
error_rates <- subset(results, select = -c(lambda, gamma))
m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
                     Avg_Error = mean(Error), Num_Reps = length(Error))
error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
error_rates <- subset(error_rates, select = -Num_Reps)

@@

@@

We used version 3.0.0 of the open-source statistical software {\tt R} for the
classification study. We used the implementations of the Witten and Guo
classifiers from the {\tt penalizedLDA} and {\tt rda} packages, respectively, on
CRAN. Because the {\tt rda} package does not perform the authors' ``Min-Min''
rule automatically, we applied this rule in our {\tt R} code. We also utilized
our implementations of the Dudoit, Pang, and Tong classifiers available in the
{\tt sparsediscrim} package. We used the default settings for each classifier
with the exception that we explicitly set prior probabilities as equal, if the
option was available. For the \emph{HDRDA} classifier, we set $\alpha_k = 1$ $(k
= 1, \ldots, K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov}
employed resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a
grid of 21 equidistant candidate values between 0 and 1, inclusively. Similarly,
we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots,
10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
10-fold cross-validation \citep{Hastie:2008dt}.

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates, fig.width=8, fig.height=8, out.width='.65\\linewidth'>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(color = Classifier), size = 1, linetype = 2)
p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 4)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 10))
p <- p + theme(axis.title.x = element_text(size = 17))
p <- p + theme(axis.title.y = element_text(size = 17))
p <- p + theme(strip.text.x = element_text(size = 16))
p <- p + theme(legend.title = element_text(size = 18, face = "bold"))
p <- p + theme(legend.text = element_text(size = 16))
p <- p + theme(legend.key.size = unit(1, "cm"))
p
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. Approximate standard errors are no
  greater than \Sexpr{max_se}.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
optimal estimates of $\lambda$ and $\gamma$ in the Supplementary Material. For
each of the four data sets, the distribution of the optimal values of $\lambda$
selected through cross-validation often had a notable bathtub shape, resembling
a Beta distribution with both shape parameters less than unity. Specifically,
the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
to quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in the \emph{HDRDA} classifier preferring the more
parsimonious assumption of equal covariance matrices. In this instance we can
see the flexibility of our proposed classifier to employ the more realistic
assumption that the class covariance matrices are unequal, but if optimal, we
can employ the more parsimonious assumption of equal covariance matrices.

@@

\section*{Supplementary material}
\label{SM}
Supplementary material available online includes the graphical summary
of the optimal estimates of our classifier from the classification study.

>>>>

\begin{aug}
\author{\fnms{John A.} \snm{Ramey}\thanksref{m1}
  \ead[label=e1]{johnramey@gmail.com}
  \ead[label=u1,url]{http://ramhiser.com}}
\author{\fnms{Caleb K.} \snm{Stein}\thanksref{m2}\ead[label=e2]{Caleb\_Stein@baylor.edu}}
\and
\author{\fnms{Dean M.} \snm{Young}\thanksref{m2}\ead[label=e3]{Dean\_Young@baylor.edu}}

@@

\address{Vaccine and Infectious Disease Division\\
  Fred Hutchinson Cancer Research Center\\
  1100 Fairview Ave. N., M2-C200\\
  Seattle, Washington 98115, U.S.A.\\
  \printead{e1}\\
  \printead{u1}}

@@

\begin{keyword}
\kwd{classification}
\kwd{regularized discriminant analysis}
\kwd{covariance-matrix regularization}
\kwd{high-dimensional microarray data}
\kwd{dimension reduction}
\end{keyword}

@@

@@

# Summary of the classification error rates for each data set and number of
# variables selected, d.
error_rates <- subset(results, select = -c(lambda, gamma))
m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
                     Avg_Error = mean(Error), Num_Reps = length(Error))
error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
error_rates <- subset(error_rates, select = -Num_Reps)
error_rates$Classifier <- factor(as.character(error_rates$Classifier))

@@

\clearpage

@@

We used version 3.0.0 of the open-source statistical software {\tt R} for the
classification study. We used the implementations of the Witten and Guo
classifiers from the {\tt penalizedLDA} and {\tt rda} packages, respectively, on
CRAN. Because the {\tt rda} package does not perform the authors' ``Min-Min''
rule automatically, we applied this rule in our {\tt R} code. We also utilized
our implementations of the Dudoit, Pang, and Tong classifiers available in the
{\tt sparsediscrim} package. We used the default settings for each classifier
with the exception that we explicitly set prior probabilities as equal, if the
option was available. For the \emph{HDRDA} classifier, we set $\alpha_k = 1$ $(k
= 1, \ldots, K)$, so that the covariance-matrix estimator \eqref{eq:hdrda-cov}
employed resembled \eqref{eq:ridge-estimator}. We estimated $\lambda$ from a
grid of 21 equidistant candidate values between 0 and 1, inclusively. Similarly,
we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots,
10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using
10-fold cross-validation \citep{Hastie:2008dt}. Each of the microarray data sets
listed below can be found in \ref{suppA}.

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates, fig.width=8, fig.height=8, out.width='.65\\linewidth'>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(color = Classifier), size = 1, linetype = 2)
p <- p + geom_point(aes(color = Classifier, shape = Classifier), size = 4)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 10))
p <- p + theme(axis.title.x = element_text(size = 17))
p <- p + theme(axis.title.y = element_text(size = 17))
p <- p + theme(strip.text.x = element_text(size = 16))
p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
p <- p + theme(legend.text = element_text(size = 15))
p <- p + theme(legend.key.size = unit(1, "cm"))
p
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. Approximate standard errors are no
  greater than \Sexpr{max_se}.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we provide graphical summaries of the
optimal estimates of $\lambda$ and $\gamma$ in \ref{suppB}. For
each of the four data sets, the distribution of the optimal values of $\lambda$
selected through cross-validation often had a notable bathtub shape, resembling
a Beta distribution with both shape parameters less than unity. Specifically,
the optimal values of $\lambda$ were typically near 0 or near 1, corresponding
to quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in the \emph{HDRDA} classifier preferring the more
parsimonious assumption of equal covariance matrices. In this instance we can
see the flexibility of our proposed classifier to employ the more realistic
assumption that the class covariance matrices are unequal, but if optimal, we
can employ the more parsimonious assumption of equal covariance matrices.

@@

\begin{supplement}
\sname{Supplement A}\label{suppA}
\stitle{R Package Containing Microarray Data Sets}
\slink[url]{https://github.com/ramey/datamicroarray}
\sdescription{An R package containing all of the microarray data sets used in
  the classification study.}
\end{supplement}

\begin{supplement}
\sname{Supplement B}\label{suppB}
\stitle{Optimal Estimates of the HDRDA Classifier}
\slink[url]{https://github.com/ramey/paper-hdrda}
\sdescription{The graphical summary of the optimal estimates of the \emph{HDRDA} classifier from the classification study.}
\end{supplement}
